---
title: "R Coding Exercise"
---

Placeholder file for the future R coding exercise.

Below we are loading the dslabs package.

```{r}

# load dslabs package
options(repos = c(CRAN = "https://github.com/cameronplayle/CameronPlayle-P2-Portfolio.git"))
install.packages("dslabs")
library(dslabs)
# load renv package
install.packages("renv")
library(renv)
# look at help file for gapminder data
help(gapminder)
# get overview of data structure 
str(gapminder)

```

```{r}

# summary of gapminder data

summary(gapminder)
```

```{r}
# checking function of type of object gapminder is

class(gapminder)
```

```{r}
# load tidyverse
library(tidyverse)
# filtering contienent variable to just Africa. I do this by using the dplyr function filter() along with pop %>%. Once this is done I can use str() and summary() to get the data structure and summaries

africadata <- gapminder %>% filter(continent == 'Africa')

str(africadata)

summary(africadata)
```

```{r}

# Now I will create two new objects, f.im and pop.gdp
# f.im will consist of the variables fertility and infant mortality 
# pop.gdp will consist of variables population and gdp

f.im <- africadata %>% select(fertility, infant_mortality)

pop.gdp <- africadata %>% select(population, gdp)

head(f.im)
head(pop.gdp)
```

```{r}
# summary of the two new objects

summary(pop.gdp)
summary(f.im)
```

```{r}

# plotting fertlilty and infant mortality 

f.im %>%
  ggplot(aes(x=fertility, y=infant_mortality)) +
  geom_point() +
  labs(x='Fertility', y='Infant Mortality',
       title='Examination of Infant Morality in African Countries',
       subtitle='Infant Morality vs Fertility')
```

```{r}

# plotting population and gdp

pop.gdp %>%
  ggplot(aes(x=population, y=gdp)) +
  coord_trans(x='log2') +
  geom_point() +
  labs(x='Population', y='Gpd',
       title='Examination of GDP in African Counties',
       subtitle='Relationship of GDP vs Population')
```

The chart imply a positive correlation between fertility and infant mortality as well as a positive correlation between population and gdp. We do see some gaps so we can isolate by grouping together population sizes. Furthermore we can group these by populations larger than 50,000

```{r}

africadata %>%
  group_by(population) %>%
  summarize(
    missing_im = sum(is.na(gdp))
  )
```

```{r}

f.im5K <- africadata %>%
  filter(population >= 50000) %>%
  select(fertility, infant_mortality)
```

```{r}

pop.gdp5K <- africadata %>%
  filter(population >= 50000) %>%
  select(population, gdp)
```

```{r}

# summary of new objects

summary(f.im5K)
summary(pop.gdp5K)
```

```{r}

# recreating process for new data

f.im5K %>%
  ggplot(aes(x=fertility, y=infant_mortality)) +
  geom_point() +
  labs(x='Fertility', y='Infant Mortality',
       title='Examination of Infant Morality in African Countries',
       subtitle='Infant Morality vs Fertility')
```

```{r}


# recreating process for new data

pop.gdp5K %>%
  ggplot(aes(x=population, y=gdp)) +
  coord_trans(x='log2') +
  geom_point() +
  labs(x='Population', y='Gpd',
       title='Examination of GDP in African Counties',
       subtitle='Relationship of GDP vs Population')
```
